\nohyphens{\textbf{Automated~Analysis~of~Weak~Memory~Models}}

### 2018-05-20 slides/slides.tex
Change the slides on encoding.

\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}

\begin{itemize}
%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
\end{itemize}

\begin{figure}[b]
  \centering
  \begin{minipage}{\textwidth}
      \begin{subfigure}[b]{0.25\textwidth}
        \makebox[\textwidth]{
        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
            \node[c] (1) [] {$e_i$};
            \node[c] (2) [below of=1] {$e_{i+1}$};
            \path[->]
            (1) edge [] node {$\epsilon$} (2)
            ;
        \end{tikzpicture}
        }
        \caption{The sequence}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.35\textwidth}
        \makebox[\textwidth] {
        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
            \node[c]           (k)   [] {$e_k$};
            \node[c,draw=none] (iii) [below of=k] {$...$};
            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
            \path[->]
            (k) edge [above] node {$g_{i,k}$} (i)
            (k) edge [] node {} (ii)
            (k) edge [] node {} (iii)
            (k) edge [] node {} (iv)
            (k) edge [above] node {$g_{i+j,k}$} (v)
            ;
        \end{tikzpicture}
        }
        \caption{Conditional branching}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.3\textwidth}
        \makebox[\textwidth]{
        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
            \node[c] (i) [] {$e_i$};
            \node[c,draw=none] (ii) [right=0.2cm of i] {};
            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
            \node[c] (k) [below of=iii] {$e_k$};
            \path[->]
            (i) edge [above] node {$g_{i,k}$} (k)
            (ii) edge [above] node {} (k)
            (iii) edge [] node {} (k)
            (iv) edge [] node {} (k)
            (v) edge [] node {} (k) %sloped
            ;
        \end{tikzpicture}
        }
        \caption{Branch merging}
    \end{subfigure}
    %
    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
    \label{encode:cf}
\end{minipage}
\end{figure}

\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}

\begin{itemize}
\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
\item The control-flow of the program is encoded as following:
\end{itemize}

\vspace{-20pt}
\begin{align}
\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
                       \land & \ \cdots \nonumber \nonumber\\
                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
\end{align}

\begin{frame}{Encoding for the data-flow}

\begin{itemize}
\item SSA-indices are computed as following:
\begin{itemize}
  \item any access to a shared variable (both read and write) increments its SSA-index;
  \item only writes to a local variable increment its SSA-index (reads preserve indices);
  \item no access to a constant variable or computed (evaluated) expression changes their SSA-index.
\end{itemize}
\end{itemize}

The data-flow of an event is encoded as following:
\begin{align}
    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \nonumber \\
    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \nonumber\\
    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] \nonumber \\
    \phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)] \nonumber
\end{align}
\end{frame}

@@

\begin{frame}{The input language}
The input language parser used by \texttt{Porthos} suffered from several disadvantages:
\begin{itemize}
\item it contained the parser code inlined directly into the grammar (hardly maintainable);
\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\lstinline{=}', `\lstinline{:=}', `\lstinline{<-}' and `\lstinline{<:-}', each for different kinds of arguments);
\item restricted syntax for expressions.
\item In contrast, \texttt{PorthosC} uses the full C language grammar of proposed in the C11 standard~\cite{jtc2011sc22} and the visitor that converts the ANTLR grammar to the AST (\texttt{Y-tree}).
\end{itemize}

@@

\begin{frame}{\texttt{X-graph} unrolling}
\begin{figure}[t]
\centering
\resizebox{.8\linewidth}{!}{
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
\node[c] (1) [] {$1$};
\node[c] (2) [below left of=1] {$2$};
\node[c] (3) [below right of=1] {$3$};
\node[c] (4) [below of=3] {$4$};
\path[->]
(1) edge [] node {} (2)
(2) edge [bend left=50,dotted] node {} (1)
(1) edge [] node {} (3)
(3) edge [] node {} (4)
(4) edge [bend right=80,dotted] node {} (1)
;
\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
;
\node[c] (22) [right=3cm of impl] {$2_2$};
\node[c] (11) [above right=1cm and 1cm of 22]{$1_1$};
\node[c] (32) [below right=1cm and 1cm of 11] {$3_2$};
\node[c] (43) [below of=32] {$4_3$};
\node[c] (13) [below of=22] {$1_3$};
\node[c] (24) [below left=1cm and 1cm of 13] {$2_4$};
\node[c] (34) [below right=1cm and 1cm of 13] {$3_4$};
\node[c] (14) [below of=43] {$1_4$};
\node[c] (15) [below of=24] {$1_5$};
\node[c] (45) [below of=34] {$4_5$};
\node[c] (25) [below of=14] {$2_5$};
\node[c] (35) [below of=14] {$3_5$};
\node[c] (26) [below left=1cm and 1cm of 15] {$2_6$};
\node[c] (36) [below right=1cm and 1cm of 15] {$3_6$};
\node[c] (16) [below right=1cm and 0.2cm of 45] {$1_6$};
\node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
\node[rectangle,draw] (6inc) [below left=1cm and 3cm of 16] {$S_{-}$};
\node[rectangle,draw] (6com) [right=3cm of 6inc] {$S_{+}$};
\node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
\node[] (k2) [right=1.7cm of 32] {$(k = 2)$};
\node[] (k3) [right=1.7cm of 43] {$(k = 3)$};
\node[] (k4) [right=1.7cm of 14] {$(k = 4)$};
\node[] (k5) [right=1.7cm of 35] {$(k = 5)$};
\node[] (k6) [right=1cm of 46] {$(k = 6)$};
\path[->]
(11) edge [] node {} (22)
(11) edge [] node {} (32)
(32) edge [] node {} (43)
(22) edge [] node {} (13)
(13) edge [bend right=10] node {} (24)
(13) edge [bend left=10] node {} (34)
(43) edge [] node {} (14)
(24) edge [] node {} (15)
(34) edge [] node {} (45)
(14) edge [] node {} (25)
(14) edge [] node {} (35)
(15) edge [bend right=10] node {} (26)
(15) edge [bend left=10] node {} (36)
(45) edge [] node {} (16)
(35) edge [bend right=10] node {} (16)
(35) edge [bend left=10] node {} (46)
(26) edge [bend right=10] node {} (6com)
(36) edge [bend right=5] node {} (6inc)
(16) edge [bend left=20] node {} (6inc)
(46) edge [bend left=20] node {} (6com)
;
\end{tikzpicture}
}
\caption{Example of the flow graph unrolling up to bound $k = 6$}
\end{figure}
\end{frame}

>>>>

%\begin{frame}{Encoding for the control-flow}% {The \texttt{PorthosC} encoding scheme}
%\begin{itemize}
%%\item The new encoding scheme of \texttt{PorthosC} follows~\cite{heljanko2008unfoldings} in general.
%\item \texttt{Porthos v1} used another encoding scheme, where the high-level instructions were represented in the SMT-formula by separate variables:
%$\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)$.
%\item In \texttt{PorthosC}, the high-level AST firstly is compiled into the \textit{event-flow graph} with events as nodes and relations as edges.
%\item All edges are labelled by \textit{guards}, local-memory computations ($\epsilon$~denotes an empty guard).
%\end{itemize}
%
%\begin{figure}[b]
%  \centering
%  \begin{minipage}{\textwidth}
%      \begin{subfigure}[b]{0.25\textwidth}
%        \makebox[\textwidth]{
%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
%            \node[c] (1) [] {$e_i$};
%            \node[c] (2) [below of=1] {$e_{i+1}$};
%            \path[->]
%            (1) edge [] node {$\epsilon$} (2)
%            ;
%        \end{tikzpicture}
%        }
%        \caption{The sequence}
%    \end{subfigure}
%    ~
%    \begin{subfigure}[b]{0.35\textwidth}
%        \makebox[\textwidth] {
%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
%            \node[c]           (k)   [] {$e_k$};
%            \node[c,draw=none] (iii) [below of=k] {$...$};
%            \node[c,draw=none] (ii)  [left=0.2cm of iii] {};
%            \node[c]           (i)   [left=0.2cm of ii] {$e_i$};
%            \node[c,draw=none] (iv)  [right=0.2cm of iii] {};
%            \node[c]           (v)   [right=0.2cm of iv] {$e_{i+j}$};
%            \path[->]
%            (k) edge [above] node {$g_{i,k}$} (i)
%            (k) edge [] node {} (ii)
%            (k) edge [] node {} (iii)
%            (k) edge [] node {} (iv)
%            (k) edge [above] node {$g_{i+j,k}$} (v)
%            ;
%        \end{tikzpicture}
%        }
%        \caption{Conditional branching}
%    \end{subfigure}
%    ~
%    \begin{subfigure}[b]{0.3\textwidth}
%        \makebox[\textwidth]{
%        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
%            \node[c] (i) [] {$e_i$};
%            \node[c,draw=none] (ii) [right=0.2cm of i] {};
%            \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
%            \node[c,draw=none] (iv) [right=0.2cm of iii] {};
%            \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
%            \node[c] (k) [below of=iii] {$e_k$};
%            \path[->]
%            (i) edge [above] node {$g_{i,k}$} (k)
%            (ii) edge [above] node {} (k)
%            (iii) edge [] node {} (k)
%            (iv) edge [] node {} (k)
%            (v) edge [] node {} (k) %sloped
%            ;
%        \end{tikzpicture}
%        }
%        \caption{Branch merging}
%    \end{subfigure}
%    %
%    \caption{Possible mutual arrangements of events in a control-flow graph} %Linear and non-linear cases of a control-flow graph}
%    \label{encode:cf}
%\end{minipage}
%\end{figure}
%\end{frame}

%\begin{frame}{Encoding for the control-flow} %{The \texttt{PorthosC} encoding scheme}
%\begin{itemize}
%\item Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted.
%\item The control-flow of the program is encoded as following:
%\end{itemize}
%\vspace{-20pt}
%\begin{align}
%\phi_{CF_{seq}} \defeq \quad & \fx(e_{i+1}) \Rightarrow \fx(e_i) \nonumber\\
%\phi_{CF_{br}}  \defeq \quad & [\fx(e_i) \Rightarrow \fx(e_k)] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \Rightarrow \fx(e_k)] \nonumber \nonumber\\
%                       \land & \ [\fx(e_i) \land \fx(e_k) \Rightarrow g_{i,k}] \ \land \ \cdots \ \land \ [\fx(e_{i+j}) \land \fx(e_k) \Rightarrow g_{i+j,k}] \nonumber \\
%                       \land & \ \cdots \nonumber \nonumber\\
%                       \land & \ ( \bigvee_{e_l \in \ \texttt{succ}(e_m)} \ \ \bigvee_{\begin{subarray}{l}e_n \in \ \texttt{succ}(e_k) \nonumber\\
%                               e_n \neq e_m\end{subarray}} \! \lnot [\fx(e_m) \land \fx(e_n)] \ ) \nonumber\\
%\phi_{CF_{mer}} \defeq \quad & \fx(e_k) \Rightarrow (\bigvee\limits_{e_p \in \ \texttt{pred}(e_k)}^{} \fx(e_p)) \nonumber
%\end{align}
%\end{frame}

\begin{frame}{Encoding for the control-flow: An example}
%Let $\fx:\,\mathbb{E}\,\rightarrow\,\{0,1\}$ be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system).
\begin{figure}
\begin{minipage}{.2\textwidth}
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
    \node[c] (i)   [] {$e_1$};
    \node[c] (ii)  [below of=i] {$e_2$};
    \node[c,fill=black!15] (lnop) [below left = 0.25cm and 0.7cm of ii] {$e_{nop_{1}}$};
    \node[c] (iii) [below of=ii] {$e_3$};
    \node[c,fill=black!15] (rnop) [right = 0.4cm of iii] {$e_{nop_{2}}$};
    \node[c] (iv)  [below of=iii] {$e_4$};
    \path[->]
    (i) edge [right] node {$\epsilon$} (ii)
    (i) edge [right,bend right=25] node {$g_{1,4}$} (lnop)
    (lnop) edge [right,bend right=25] node {$\epsilon$} (iv)
    (ii) edge [right] node {$g_{2,3}$} (iii)
    (iii) edge [right] node {$\epsilon$} (iv)
    (ii) edge [right,bend left=25] node {$g_{2,4}$} (rnop)
    (rnop) edge [right,bend left=25] node {$\epsilon$} (iv)
    ;
\end{tikzpicture}
\end{minipage}
%
\hspace{15pt}
\begin{minipage}{.4\textwidth}
\centering
\begin{align*}
\phi_{CF} \defeq \ & [\fx(e_2) \Rightarrow \fx(e_1)] \\
                 \ & \land [\fx(e_3) \Rightarrow \fx(e_2)] \\
                 \ & \land [\fx(e_{nop_{1}}) \Rightarrow \fx(e_1)] \\
                 \ & \land [\fx(e_{nop_{2}}) \Rightarrow \fx(e_2)] \\
                 \ & \land [\fx(e_4) \Rightarrow (\fx(e_{nop_{1}}) \lor \fx(e_3) \lor \fx(e_{nop_{2}}))] \\
                 \ & \land [\fx(e_{nop_{1}}) \land \fx(e_1) \Rightarrow g_{1,4}] \\
                 \ & \land [(\fx(e_3) \land \fx(e_2)) \Rightarrow g_{2,3}] \\
                 \ & \land [(\fx(e_{nop_{2}}) \land \fx(e_2)) \Rightarrow g_{2,4}] \\
                 \ & \land \lnot [\fx(e_2) \land \fx(e_{nop_{1}})] \\
                 \ & \land \lnot [\fx(e_3) \land \fx(e_{nop_{2}})]
\end{align*}
\end{minipage}
\caption{Example of encoding for the control-flow of the event-flow graph}
\end{figure}
\end{frame}

\begin{frame}{Encoding for the data-flow}
\begin{itemize}
\item SSA-indices are computed as following:
\begin{itemize}
  \item any access to a shared variable (both read and write) increments its SSA-index;
  \item only writes to a local variable increment its SSA-index (reads preserve indices);
  \item no access to a constant variable or computed (evaluated) expression changes their SSA-index.
\end{itemize}
\end{itemize}
The data-flow of an event is encoded as following:
\begin{align}
    \phi_{DF_{e = \texttt{load}(r \leftarrow l)}}  \defeq \ & [\fx(e) \Rightarrow (r_{i+1} = l_{i+1})] \nonumber \\
    \phi_{DF_{e = \texttt{store}(l \leftarrow r)}} \defeq \ & [\fx(e) \Rightarrow (l_{i+1} = r_i)] \nonumber\\
    \phi_{DF_{e = \texttt{eval}(\cdot)}}           \defeq \ & [\fx(e) \Rightarrow \fv(e)] \nonumber \\
\end{align}
%    \phi_{DF_{mem}}(e_1, e_2) \defeq \ & [\texttt{rf}(e_1, e_2) \Rightarrow (l_i = l_j)] \nonumber
\end{frame}

@@

\begin{frame}{The input language}
The input language parser used by \texttt{Porthos} suffered from several disadvantages:
\begin{itemize}
\item it contained the parser code inlined directly into the grammar (hardly maintainable);
\item the semantics of operations and kinds of variables (global or shared) were determined syntactically (4 different types of assignment:`\texttt{=}', `\texttt{:=}', `\texttt{<-}' and `\texttt{<:-}', each for different kinds of arguments);
\item restricted syntax for expressions.
\item In contrast, \texttt{PorthosC} uses the full C language grammar of proposed in the C11 standard~\cite{jtc2011sc22} and the visitor that converts the ANTLR grammar to the AST (\texttt{Y-tree}).
\end{itemize}

@@

\begin{frame}{\texttt{X-graph} unrolling}
\begin{figure}[t]
\centering
\resizebox{.8\linewidth}{!}{
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
\node[c] (1) [] {$1$};
\node[c] (2) [below left of=1] {$2$};
\node[c] (3) [below right of=1] {$3$};
\node[c] (4) [below of=3] {$4$};
\path[->]
(1) edge [] node {} (2)
(2) edge [bend left=50,dotted] node {} (1)
(1) edge [] node {} (3)
(3) edge [] node {} (4)
(4) edge [bend right=80,dotted] node {} (1)
;
\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
;
\node[c] (22) [right=3cm of impl] {$2_2$};
\node[c] (11) [above right=1cm and 1cm of 22]{$1_1$};
\node[c] (32) [below right=1cm and 1cm of 11] {$3_2$};
\node[c] (43) [below of=32] {$4_3$};
\node[c] (13) [below of=22] {$1_3$};
\node[c] (24) [below left=1cm and 1cm of 13] {$2_4$};
\node[c] (34) [below right=1cm and 1cm of 13] {$3_4$};
\node[c] (14) [below of=43] {$1_4$};
\node[c] (15) [below of=24] {$1_5$};
\node[c] (45) [below of=34] {$4_5$};
\node[c] (25) [below of=14] {$2_5$};
\node[c] (35) [below of=14] {$3_5$};
\node[c] (26) [below left=1cm and 1cm of 15] {$2_6$};
\node[c] (36) [below right=1cm and 1cm of 15] {$3_6$};
\node[c] (16) [below right=1cm and 0.2cm of 45] {$1_6$};
\node[c] (46) [below right=1cm and 0.3cm of 35] {$4_6$};
\node[rectangle,draw] (6) [below left=1cm and 1cm of 16] {$S$};
\node[] (k1) [right=3.2cm of 11] {$(k = 1)$};
\node[] (k2) [right=1.7cm of 32] {$(k = 2)$};
\node[] (k3) [right=1.7cm of 43] {$(k = 3)$};
\node[] (k4) [right=1.7cm of 14] {$(k = 4)$};
\node[] (k5) [right=1.7cm of 35] {$(k = 5)$};
\node[] (k6) [right=1cm of 46] {$(k = 6)$};
\path[->]
(11) edge [] node {} (22)
(11) edge [] node {} (32)
(32) edge [] node {} (43)
(22) edge [] node {} (13)
(13) edge [bend right=10] node {} (24)
(13) edge [bend left=10] node {} (34)
(43) edge [] node {} (14)
(24) edge [] node {} (15)
(34) edge [] node {} (45)
(14) edge [] node {} (25)
(14) edge [] node {} (35)
(15) edge [bend right=10] node {} (26)
(15) edge [bend left=10] node {} (36)
(45) edge [] node {} (16)
(35) edge [bend right=10] node {} (16)
(35) edge [bend left=10] node {} (46)
(26) edge [bend right=10] node {} (6)
(36) edge [bend right=5] node {} (6)
(16) edge [bend left=20] node {} (6)
(46) edge [bend left=20] node {} (6)
;
\end{tikzpicture}
}
\caption{Example of the flow graph unrolling up to bound $k = 6$}
\end{figure}
\end{frame}
